---
title: "Running Analytics"
author: "Jacob Duvall"
date: "Originally published February 1, 2019"
output: html_document
runtime: shiny
---

<style type="text/css"> 

h1.title { 
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>

# https://community.rstudio.com/t/embed-a-shiny-app-in-a-private-website/19163/4
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This application uses publicly available data to showcase running analytics over some popular distances.
```{r eruptions, echo=FALSE}
library(ggplot2) #the graphs
library(lubridate) #mdy
library(reshape2) #melt
library(shiny) #shinyApp
# library(plyr)
# library(gridExtra)

#data import men's marathon and women's marathon
marathon.df = read.csv("bothMarathon.csv")
marathon.df1 = marathon.df
marathon = subset(marathon.df, GENDER == 'Male')
wmarathon = subset(marathon.df, GENDER == 'Female')

#data import men's half marathon and women's half marathon
halfMarathon.df = read.csv("halfMarathon.csv")
halfMarathon = subset(halfMarathon.df, GENDER == 'MALE')
whalfMarathon = subset(halfMarathon.df, GENDER == 'FEMALE')

#set marathon dates in decreasing order
marathon$DATE = mdy(marathon$DATE)
dq = marathon[order(marathon$DATE,decreasing =T ),]
wmarathon$DATE = mdy(wmarathon$DATE)
dq = wmarathon[order(wmarathon$DATE,decreasing=T),]
marathon.df$DATE = mdy(marathon.df$DATE)
dq = marathon.df[order(marathon.df$DATE,decreasing =T ),]

#set half marathon dates in decreasing order
halfMarathon$DATE = mdy(halfMarathon$DATE)
dq = halfMarathon[order(halfMarathon$DATE,decreasing =T ),]
whalfMarathon$DATE = mdy(whalfMarathon$DATE)
dq = whalfMarathon[order(whalfMarathon$DATE,decreasing=T),]
halfMarathon.df$DATE = mdy(halfMarathon.df$DATE)
dq = halfMarathon.df[order(halfMarathon.df$DATE,decreasing =T ),]

#convert marathon string times into numerics 
x = paste(marathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
marathon$TIME = x 
 
x = paste(wmarathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
wmarathon$TIME = x 
 
x = paste(marathon.df$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
marathon.df$TIME = x 

#convert half marathon string times into numerics 
x = paste(halfMarathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
halfMarathon$TIME = x 

x = paste(whalfMarathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
whalfMarathon$TIME = x 

x = paste(halfMarathon.df$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
halfMarathon.df$TIME = x 

marathon
# Input options
inputPanel(
  selectInput("plotg", "Choose Race Option:",
      list(`options` = c("Marathon", "Half Marathon")
      )
    ),
  selectInput("rd1", "Select Gender:",
              list(`options` = c("Male", "Female", "Both")
              )
              ),
  checkboxInput("table", label = "Show Table", value = TRUE))

# plot renderings 
renderPlot({
  
# ---------------------  MARATHON PLOTS ---------------------
# men's marathon plot  
marathonGraph = ggplot(data = marathon, aes(x = DATE, y = TIME)) +  geom_line(color = "steelblue") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's Half Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 

# women's marathon plot
wmarathonGraph = ggplot(data = wmarathon, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "deeppink") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Women's Marathon Record Over Time", x= "DATE", y ="FINISHING TIME")

# both men's and women's marathon plot
bothmarathonGraph = ggplot(data = marathon.df, aes(x=DATE, y =TIME,col=GENDER)) + geom_line() + geom_point() + theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's and Women's Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 

# --------------------- HALF MARATHON PLOTS ---------------------

# men's half marathon plot
halfmarathonGraph = ggplot(data = halfMarathon, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "steelblue") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's Half Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 

# women's half marathon plot
whalfmarathonGraph = ggplot(data = whalfMarathon, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "deeppink") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Womens's Half Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 

# both men's and women's half marathon plot
bothhalfmarathon = ggplot(data = halfMarathon.df, aes(x=DATE, y =TIME,col=GENDER)) + geom_line() + geom_point() + theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's and Women's Half Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 


# men's and women's marathon prints
 if(input$plotg == "Marathon" & input$rd1 == "Male") print(marathonGraph) 
 if(input$plotg == "Marathon" & input$rd1 == "Female") print(wmarathonGraph)
 if(input$plotg == "Marathon" & input$rd1 == "Both") print(bothmarathonGraph)


#men's and women's half marathon prints
 if(input$plotg == "Half Marathon" & input$rd1 == "Male") print(halfmarathonGraph) 
 if(input$plotg == "Half Marathon" & input$rd1 == "Female") print(whalfmarathonGraph)
 if(input$plotg == "Half Marathon" & input$rd1 == "Both") print(bothhalfmarathon)
})

  shinyApp(
    
       ui = fluidPage(
          fluidRow(
           column(12,
             dataTableOutput('table')
           )
         )
       ),
      server = function(input, output) {
                 output$table <- renderDataTable({
           {
               if(!is.null(input$rd2 == 100)) {
                   marathon.df1
               }
             else {
               NULL
             }
           }})
         }
       )



```
#bugs - y axis is 1 unit each entry

