---
title: "Running Analytics"
author: "Jacob Duvall"
date: "Originally published February 1, 2019"
output: html_document
runtime: shiny
---

<style type="text/css"> 

h1.title { 
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>

# https://community.rstudio.com/t/embed-a-shiny-app-in-a-private-website/19163/4
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This application uses publicly available data to showcase running analytics over some popular distances.
```{r eruptions, echo=FALSE}
library(ggplot2) #the graphs
library(lubridate) #mdy
#library(ggiraph) # interactive graph

#data import men's marathon and women's marathon
marathon.df = read.csv("bothMarathon.csv")
marathon.df1 = marathon.df
marathon = subset(marathon.df, GENDER == 'Male')
marathon.df2 = marathon
wmarathon = subset(marathon.df, GENDER == 'Female')
wmarathon.df = wmarathon

#data import men's half marathon and women's half marathon
halfMarathon.df = read.csv("halfMarathon.csv")
halfMarathon.df1 = halfMarathon.df
halfMarathon = subset(halfMarathon.df, GENDER == 'MALE')
halfMarathon.df2 = halfMarathon
whalfMarathon = subset(halfMarathon.df, GENDER == 'FEMALE')
whalfMarathon.df = whalfMarathon

#data import men's 10k and women's 10k
tenk.df = read.csv("10k.csv")
tenk.df1 = tenk.df
tenk = subset(tenk.df, GENDER == 'MALE')
tenk.df2 = tenk
wtenk = subset(tenk.df, GENDER == 'FEMALE')
wtenk.df = wtenk

#set marathon dates in decreasing order
marathon$DATE = mdy(marathon$DATE)
dq = marathon[order(marathon$DATE,decreasing =T ),]
wmarathon$DATE = mdy(wmarathon$DATE)
dq = wmarathon[order(wmarathon$DATE,decreasing=T),]
marathon.df$DATE = mdy(marathon.df$DATE)
dq = marathon.df[order(marathon.df$DATE,decreasing =T ),]

#set half marathon dates in decreasing order
halfMarathon$DATE = mdy(halfMarathon$DATE)
dq = halfMarathon[order(halfMarathon$DATE,decreasing =T ),]
whalfMarathon$DATE = mdy(whalfMarathon$DATE)
dq = whalfMarathon[order(whalfMarathon$DATE,decreasing=T),]
halfMarathon.df$DATE = mdy(halfMarathon.df$DATE)
dq = halfMarathon.df[order(halfMarathon.df$DATE,decreasing =T ),]

#set 10k dates in decreasing order
tenk$DATE = mdy(tenk$DATE)
dq = tenk[order(tenk$DATE,decreasing =T ),]
wtenk$DATE = mdy(wtenk$DATE)
dq = wtenk[order(wtenk$DATE,decreasing=T),]
tenk.df$DATE = mdy(tenk.df$DATE)
dq = tenk.df[order(tenk.df$DATE,decreasing =T ),]

#convert marathon string times into numerics 
x = paste(marathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
marathon$TIME = x 
 
x = paste(wmarathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
wmarathon$TIME = x 
 
x = paste(marathon.df$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
marathon.df$TIME = x 

#convert half marathon string times into numerics 
x = paste(halfMarathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
halfMarathon$TIME = x 

x = paste(whalfMarathon$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
whalfMarathon$TIME = x 

x = paste(halfMarathon.df$TIME)
x = as.POSIXct(strptime(x, "%H:%M:%S"))
halfMarathon.df$TIME = x 

#convert 10k string times into numerics 
x = paste(tenk$TIME)
x = as.POSIXct(strptime(x, "%M:%S"))
tenk$TIME = x 

x = paste(wtenk$TIME)
x = as.POSIXct(strptime(x, "%M:%S"))
wtenk$TIME = x 

x = paste(tenk.df$TIME)
x = as.POSIXct(strptime(x, "%M:%S"))
tenk.df$TIME = x 

# Input options
inputPanel(
  selectInput("plotg", "Choose Race Option:",
      list(`options` = c("Marathon", "Half Marathon", "10,000 Meters", "5,000 Meters", "1 Mile")
      )
    ),
  selectInput("rd1", "Select Gender:",
              list(`options` = c("Male", "Female", "Both")
              )
              ),
  checkboxInput("table", label = "Show Data Table", value = TRUE),
  selectInput("additional", "Show Additional Analytics:",
              list(`options` = c("Nationality Bar Graph", "Nationality Pie Chart"))))

# -------------------- RUNNING PLOTS -----------------------

renderPlot({
  
# ---------------------  MARATHON PLOTS ---------------------
# men's marathon plot  
marathonGraph = ggplot(data = marathon, aes(x = DATE, y = TIME)) +  geom_line(color = "steelblue") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's Half Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 


# women's marathon plot
wmarathonGraph = ggplot(data = wmarathon, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "deeppink") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Women's Marathon Record Over Time", x= "DATE", y ="FINISHING TIME")

# both men's and women's marathon plot
bothmarathonGraph = ggplot(data = marathon.df, aes(x=DATE, y =TIME,col=GENDER)) + geom_line() + geom_point() + theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's and Women's Marathon Records Over Time", x= "DATE", y ="FINISHING TIME") 

# --------------------- HALF MARATHON PLOTS ---------------------

# men's half marathon plot
halfmarathonGraph = ggplot(data = halfMarathon, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "steelblue") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's Half Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 

# women's half marathon plot
whalfmarathonGraph = ggplot(data = whalfMarathon, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "deeppink") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Womens's Half Marathon Record Over Time", x= "DATE", y ="FINISHING TIME") 

# both men's and women's half marathon plot
bothhalfmarathon = ggplot(data = halfMarathon.df, aes(x=DATE, y =TIME,col=GENDER)) + geom_line() + geom_point() + theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's and Women's Half Marathon Records Over Time", x= "DATE", y ="FINISHING TIME") 

# ----------------- 10K PLOTS ----------------------

tenkGraph = ggplot(data = tenk, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "steelblue") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's 10K Record Over Time", x= "DATE", y ="FINISHING TIME") 

# women's half marathon plot
wtenkGraph = ggplot(data = wtenk, aes(x = DATE, y = TIME, group = 1)) +  geom_line(color = "deeppink") + geom_point() +
  theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Womens's 10K Record Over Time", x= "DATE", y ="FINISHING TIME") 

# both men's and women's half marathon plot
bothtenkGraph = ggplot(data = tenk.df, aes(x=DATE, y =TIME,col=GENDER)) + geom_line() + geom_point() + theme(axis.text.x = element_text(angle=70, vjust=0.6))  +
  labs(title="Men's and Women's 10K Records Over Time", x= "DATE", y ="FINISHING TIME") 

# ----------------- PRINT PLOTS --------------------

# men's and women's marathon prints
 if(input$plotg == "Marathon" & input$rd1 == "Male") print(marathonGraph)
 if(input$plotg == "Marathon" & input$rd1 == "Female") print(wmarathonGraph)
 if(input$plotg == "Marathon" & input$rd1 == "Both") print(bothmarathonGraph)


# men's and women's half marathon prints
 if(input$plotg == "Half Marathon" & input$rd1 == "Male") print(halfmarathonGraph) 
 if(input$plotg == "Half Marathon" & input$rd1 == "Female") print(whalfmarathonGraph)
 if(input$plotg == "Half Marathon" & input$rd1 == "Both") print(bothhalfmarathon)

# men's and women's 10K prints
 if(input$plotg == "10,000 Meters" & input$rd1 == "Male") print(tenkGraph) 
 if(input$plotg == "10,000 Meters" & input$rd1 == "Female") print(wtenkGraph)
 if(input$plotg == "10,000 Meters" & input$rd1 == "Both") print(bothtenkGraph)
})

# ------------------------ ADDITIONAL ANALYTICS ------------------


# ------------------------ NATIONALITY BAR GRAPH/PIE CHART -------

renderPlot({
  
if(input$plotg == "Marathon" & input$rd1 == "Male" & input$additional == "Nationality Bar Graph") {
  x = barplot(table(marathon$COUNTRY), xaxt="n",col="steelblue",main="Number of Marathon Records Per Country")
  labs = paste(names(table(marathon$COUNTRY)), "")
  text(cex=.9, x=x-0, y=-1, labs, xpd=TRUE, srt=75)
}

if(input$plotg == "Marathon" & input$rd1 == "Male" & input$additional == "Nationality Pie Chart") {
  pie(with(marathon,table(COUNTRY)),col=1:14,main = "Number of Marathon Records Per Country")
}
  
})

# ------------------------ DATA TABLES ---------------------------
# Mens's Marathon table
renderDataTable({
 if(input$plotg == "Marathon" & input$rd1 == "Male" & input$table == TRUE) marathon.df2
  })

# Women's Marathon table
renderDataTable({
 if(input$plotg == "Marathon" & input$rd1 == "Female" & input$table == TRUE) wmarathon.df
  })

# Men's and Women's Marathon table
renderDataTable({
 if(input$plotg == "Marathon" & input$rd1 == "Both" & input$table == TRUE) marathon.df1
  })

# Mens's Half Marathon table
renderDataTable({
 if(input$plotg == "Half Marathon" & input$rd1 == "Male" & input$table == TRUE) halfMarathon.df2
  })

# Women's Half Marathon table
renderDataTable({
 if(input$plotg == "Half Marathon" & input$rd1 == "Female" & input$table == TRUE) whalfMarathon.df
  })

# Men's and Women's Half Marathon table
renderDataTable({
 if(input$plotg == "Half Marathon" & input$rd1 == "Both" & input$table == TRUE) halfMarathon.df1
  })

# Mens's 10K table
renderDataTable({
 if(input$plotg == "10,000 Meters" & input$rd1 == "Male" & input$table == TRUE) tenk.df2
  })

# Women's 10K table
renderDataTable({
 if(input$plotg == "10,000 Meters" & input$rd1 == "Female" & input$table == TRUE) wtenk.df
  })

# Men's and Women's 10K table
renderDataTable({
 if(input$plotg == "10,000 Meters" & input$rd1 == "Both" & input$table == TRUE) tenk.df1
  })


```
renderPlot({
  marathonGraph = ggplot(marathon, aes(x= DATE, y = TIME, color = GENDER))
  marathonGraph1 = marathonGraph + geom_point_interactive(aes(tooltip = ATHLETE), size = 2)
  
  girafe(code = print(marathonGraph1))
  })
